Searching for long-period transiting planets in the \kepler\ light curves using supervised classification

### 2015-04-08 document/ms.tex
words about training

\newcommand{\BIC}{{\ensuremath{\mathrm{BIC}}}}
\newcommand{\TIC}{{\ensuremath{\mathrm{TIC}}}}
\newcommand{\T}{\ensuremath{\mathrm{T}}}
\newcommand{\dd}{\ensuremath{\,\mathrm{d}}}
\newcommand{\bvec}[1]{{\ensuremath{\boldsymbol{#1}}}}
\newcommand{\appropto}{\mathrel{\vcenter{
  \offinterlineskip\halign{\hfil$##$\cr
    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
\newcommand{\densityunit}{{\ensuremath{\mathrm{nat}^{-2}}}}

@@

% Notation for this paper.
\newcommand{\period}{{\ensuremath{P}}}
\newcommand{\rp}{{\ensuremath{R_\mathrm{P}}}}
\newcommand{\rate}{{\ensuremath{\Gamma}}}

@@

Nearly every transit search algorithm is built on the same principles and many
of the same decisions are made.
In particular, at the heart of most methods is a matched filtering step where
the likelihood of an approximate transit model is computed on a grid in the
physical parameters \dfmtodo{cite}.
Using these methods, and substantial hand-curation, some long-period
transiting candidates have been published \citep[for example][]{Batalha:2013,
Huang:2013, Kipping:2014a}.
For more recent data releases, the community has settled on more conservative
selection criteria where candidates are required to have multiple transits
\citep[for example][]{Petigura:2013, Burke:2014}.
This means that there has never been a systematic search for single transits
in the full \kepler\ dataset.

@@

The \kepler\ Mission measured photometry for about 190,000 stars at half-hour
cadence for a baseline of over four years.
We aim to search these time series for single transits of long-period planets
and single eclipses of binary stars.
These data are made available on
MAST\footnote{\url{https://archive.stsci.edu/kepler/}} and, for each target,
we downloaded the full set of long cadence light curve files provided by Data
Release 24 \citep{Thompson:2015}.
From these files, we extracted the PDC time series and split them into
``sections'' with no more than ten contiguous missing or flagged data points.
The PDC light curves have been corrected for the instrumental effects caused
by the spacecraft using a data-driven model of the focal plane
\citep{Stumpe:2012, Smith:2012}.
Crucially, an attempt is also made by the PDC procedure to remove sharp
instrumental artifacts like ``sudden pixel sensitivity dropouts (SPSDs)''.
The success rate of this correction procedure is much higher than in earlier
data releases but, as discussed in \sect{demo}, there remain some cases that
are not properly accounted for.

@@

\paragraph{Terminology}

Any remaining missing data points are filled in using linear interpolation
resulting in sections of contiguous time series measurements with nearly
uniform time sampling.
This procedure yields about 100 \emph{light curve sections} for the typical
\kepler\ target.

As we discuss further below, a key assumption of our analysis is that the
process generating any non-transit variability or noise in the light curve
is stationary.

Every three months, the \kepler\ spacecraft rolled 90~degrees to maintain its
orientation with respect to the Sun.
This means that each star landed on a different set of pixels each quarter
and, therefore, the systematic signals caused by any specific spacecraft
orientation won't be shared between neighboring quarters.
A year later, however, the pointing would return to nearly the same setup and
each star ends up back on many of the same pixels.
This manifests itself in the light curves by causing the observations
separated by a year to be qualitatively similar.

For reasons that will become clear in \sect{search}, we split the light curve
sections into 3 disjoint sets.
We use these sets to run a train/validate/test procedure so the goal is to
have representative samples in each set.
Therefore, we attempt to split the sections intelligently; evenly
distributing sections from the same quarter and season across the splits.
When this is not possible, we randomly assign the section to a split.

@@

\subsection{Splitting the dataset}

It is crucial that the model used to test for transits in a raw \kepler\
light curve not be trained on the same data.
We must split the light curve of each star into 3 disjoint sets.
For the model to work well, we require that the signals induced by
variability, noise, and other systematics be consistent between splits.
It is, therefore, necessary that we assign the splits carefully.
As discussed in \sect{data}, a light curve can be separated naturally into
``sections'' of contiguous measurements; these are the base unit for splitting
the dataset.
For the typical \kepler\ target, this division is finer than month long
chunks and there are about 100 sections for each light curve.
Each light curve section is cut out of a specific ``quarter'' and each quarter
was observed during a specific ``season''.

@@

\subsection{Initial candidate detection}

>>>>

\newcommand{\T}{\ensuremath{\mathrm{T}}}
\newcommand{\dd}{\ensuremath{\,\mathrm{d}}}
\newcommand{\bvec}[1]{{\ensuremath{\boldsymbol{#1}}}}
\newcommand{\appropto}{\mathrel{\vcenter{
  \offinterlineskip\halign{\hfil$##$\cr
    \propto\cr\noalign{\kern2pt}\sim\cr\noalign{\kern-2pt}}}}}
\newcommand{\densityunit}{{\ensuremath{\mathrm{nat}^{-2}}}}

@@

% Notation for this paper.
\newcommand{\period}{{\ensuremath{P}}}
\newcommand{\rp}{{\ensuremath{R_\mathrm{P}}}}
\newcommand{\rate}{{\ensuremath{\Gamma}}}
\newcommand{\score}{{\ensuremath{S}}}

@@

Nearly every transit search algorithm is built on the same principles and many
of the same decisions are made.
In particular, at the heart of most methods is a matched filtering step where
the likelihood of an approximate transit model is computed on a grid in the
physical parameters \dfmtodo{cite}.
Using these methods, and substantial hand-curation, some long-period
transiting candidates have been published \citep[for example][]{Batalha:2013,
Huang:2013, Kipping:2014a}.
For more recent data releases, the community has settled on more conservative
selection criteria where candidates are required to have multiple transits
\citep[for example][]{Petigura:2013, Burke:2014}.
Even the \project{QATS} algorithm \dfmtodo{cite} for finding quasiperiodic
transits builds on much the same infrastructure.
This means that there has never been a systematic search for single transits
in the full \kepler\ dataset.

@@

The \kepler\ Mission measured photometric time series for about 190,000 stars
at half-hour cadence for a baseline of over four years.
We aim to search these light curves for single transits of long-period planets
and single eclipses of binary stars.
These data are made available on
MAST\footnote{\url{https://archive.stsci.edu/kepler/}} and, for each target,
we downloaded the full set of long cadence light curve files provided by Data
Release 24 \citep{Thompson:2015}.
From these files, we extracted the PDC time series and split them into
``sections'' with no more than ten contiguous missing or flagged data points.
The PDC light curves have been corrected for the instrumental effects caused
by the spacecraft using a data-driven model of the focal plane
\citep{Stumpe:2012, Smith:2012}.
Crucially, an attempt is also made by the PDC procedure to remove sharp
instrumental artifacts like ``sudden pixel sensitivity dropouts (SPSDs)''.
The success rate of this correction procedure is much higher than in earlier
data releases but, as discussed in \sect{demo}, there remain some cases that
are not properly accounted for.

@@

@@

\subsection{Splitting the dataset}\sectlabel{split}

It is crucial that the model used to test for transits in a raw \kepler\
light curve not be trained on the same data.
We must split the light curve of each star into 3 disjoint sets.
For the model to work well, we require that the signals induced by
variability, noise, and other systematics be consistent between splits.
It is, therefore, necessary that we assign the splits carefully.
As discussed in \sect{data}, a light curve can be separated naturally into
``sections'' of nearly contiguous measurements; these are the base unit for
splitting the dataset.
For the typical \kepler\ target, this division is finer than month long
chunks and there are about 100 sections for each light curve.
Each light curve section is cut out of a specific ``quarter'' and each quarter
was observed during a specific ``season''.

@@

The dataset will still have some missing or flagged fluxes.
In order to account for these missing points, we linearly interpolate these
values based on their neighbors.
For consistency, we perform the interpolation on the training set \emph{after
injecting the synthetic transit signals}.
Before passing the light curve sections to the model at any stage, we
normalize the section by its empirical median and take the logarithm of the
fluxes.
This normalization appears to yield better performance in some test cases but
better performance might be achieved using optimized methods.
This choice is discussed further in \sect{params}.

\subsection{Training, validation, and testing}

For each subset of the data, we train a RF classifier on the 40,000 training
examples generated from the light curves in that split.
The \texttt{RandomForestClassifier} implementation from \project{scikit-learn}
has several tuning parameters that could be optimized using cross-validation
but we find acceptable performance in most cases using fixed values.
Specifically, we use a forest with 1000 trees by setting
\texttt{num\_estimators = 1000}.
We discuss this choice and the settings for other parameters in more detail in
\sect{params}.

Using this classifier, fit to one of the light curve subsets, we validate its
performance on each set of simulations from the two other splits.
This results in two precision--recall curves for the classifier.
The \emph{precision} is an empirical measurement of the sample purity (or
false positive rate) and \emph{recall} is a measurement of the completeness
(the probability of detecting a true signal).
Since the RF classifier scores each class and with a continuous value \score\
between 0 and 1, both precision and recall can be tuned by changing the
threshold value of \score\ above which the signal is considered a candidate.
For each model--validation set pair, we ambitiously choose the threshold in
order to obtain a precision of 100~percent.
The result of this procedure applied to the three splits is three classifiers
with two score thresholds each.

Each of these six models has been fit using two splits so we then apply it to
the raw light curves (without injected signals) from the remaining, untouched
splits.
This yields a \texttt{transit} class score as a function of transit time and
any points above the detection threshold are passed along to the candidate
selection procedure.

At this point, there are two class predictions from two different models for
every light curve section.
Any section that is classified as \texttt{transit} by \emph{both models} is
accepted as a single transit candidate.

\dfmtodo{make a figure}
